A Method for Regularisation of Evolutionary Polynomial Regressions using Support Vector Machines

### 2013-11-20 texts/submission_02-asc/paper_spfga-review.tex
Completed introduction.

With notable exceptions (\emph{e.g.} neural networks) machine learning regression techniques produce linear models. The linearity assumption has many advantages including reduced computational complexity and strong the\-o\-re\-ti\-cal framework. However nonlinearity is unavoidable in many application scenarios, specially those with phase transitions or feedback loops, so common in engineering, ecology, cybernetics and other areas. The kernel trick in \acp{SVM} (\cite{scholkopf1997kernel,liang2012eigen,Bao:2013aa}) alleviates this problem by allowing special non-linear transformations of the feature-space. The condition such transformations must meet is known as the \emph{kernel trick}, $k\at{x,x'} = \left< \varphi\at{x}, \varphi\at{x'} \right>$, where $\varphi$ is the feature-space transformation and $\left<\cdot,\cdot\right>$ denotes the inner product. The ``trick'' consists on computing the kernel $k\at{x,x'}$ while avoiding the computation of the inner product and the transformations $\varphi\at{x}, \varphi\at{x'}$. A special case of polynomial transformation, the \emph{polynomial kernel}, $k\at{x,x'} = \left<x,x'\right>^d$ is commonly used in regression and classification tasks with \acp{SVM}. However general polynomial transformations do not verify the kernel trick.

Polynomials, one of the most studied subjects in mathematics, generalise li\-ne\-ar functions and define, perhaps, the simplest and most used nonlinear models. Applications include colorimetric calibration \citep{Mendes:2005aa}, explicit formul\ae\ for turbulent pipe flows
 \citep{Davidson:1999aa}, computational linguistics \citep{Sanchez:2009aa} and more recently analytical techniques for cultural heritage materials \citep{Csefalvayova:2010aa}, liquid epoxy moulding process \citep{Chan:2011aa}, B-spline surface reconstruction \citep{Galvez:2012aa}, product design \citep{Chan:2012aa} or forecasting cyanotoxins presence in water reservoirs \citep{Garcia-Nieto:2013aa}. These examples not only illustrate the wide spectrum of applications but, additionally, work in each one uses, at some point, a genetic algorithm.

\ac{GA} where, arguably, one of the hottest topics of research in the recent decades and with good reason since they outline an optimisation scheme easy to conceptualise and with very broad application. If a nonlinear (or otherwise) model requires parameterization \acp{GA} provide a simple and often effective approach to search for locally optimal parameters. Research related to genetic algorithms abound and spans from the 1950s seminal work of Nils Aall Barricelli \citep{Barricelli:1962aa} in the Institute for Advanced Study of Princeton to today's principal area of study for thousands of researchers, covered in hundreds of conferences, workshops and other meetings. Perhaps the key impulse to \acp{GA} come from John Holland's work and his book ``Adaptation in Natural and Artificial Systems'' \citep{Holland:1975aa}.

One interesting take on genetic algorithms, named \emph{genetic programming} by John Koza \citep{Koza:1992aa}, proposed the use of \acp{GA} to search the syntactic structure of complex functions. This syntactic structure search is keen to the central ideas of deep learning \citep{Bengio:2009aa,Bengio:2013aa}, a subarea of machine learning actually producing quite promising results (\emph{e.g.} in \cite{Tarlow:2013fk}). It is also related to the work presented in this paper in the sense that, unlike linear models that have a simple structure, $y=\sum_i \beta_i x_i$, nonlinear (in particular polynomial) models pose an additional ``structure" search problem.

@@

\note{Redo the purpose of this work.} In line with that research this work describes a general method to find a polynomial regression of a given dataset. The optimal regression minimizes a cost function that accounts for both the \ac{rmse} and a regularization factor to avoid over-fitting.

@@

It turns out that, discarding the computational cost of training, the polynomial regression method presented here, \acf{EPRR}, provides a quite effective regression method. Indeed, according to the results of common regression methods on several real-world datasets it is only systematically out-performed by random forests, an \emph{ensemble} method.

>>>>

With notable exceptions (\emph{e.g.} neural networks) machine learning regression techniques produce linear models. The linearity assumption has many advantages including reduced computational complexity and strong the\-o\-re\-ti\-cal framework. However nonlinearity is unavoidable in many application scenarios, specially those with phase transitions or feedback loops, so common in engineering, ecology, cybernetics and other areas. The kernel trick in \acp{SVM} (\cite{scholkopf1997kernel,liang2012eigen,Bao:2013aa}) alleviates this problem by allowing special non-linear transformations of the feature-space. The condition such transformations must meet is known as the \emph{kernel trick}, $k\at{x,x'} = \left< \varphi\at{x}, \varphi\at{x'} \right>$, where $\varphi$ is the feature-space transformation and $\left<\cdot,\cdot\right>$ denotes inner product. The ``trick'' consists on computing the kernel $k\at{x,x'}$ while avoiding the computation of the inner product and the transformations $\varphi\at{x}, \varphi\at{x'}$. A special case of polynomial transformation, the \emph{polynomial kernel}, $k\at{x,x'} = \left<x,x'\right>^d$ is commonly used in regression and classification tasks with \acp{SVM}. However general polynomial transformations do not verify the kernel trick.

Polynomials, one of the most studied subjects in mathematics, generalise li\-ne\-ar functions and define, perhaps, the simplest and most used nonlinear models. Applications include colorimetric calibration \citep{Mendes:2005aa}, explicit formul\ae\ for turbulent pipe flows
 \citep{Davidson:1999aa}, computational linguistics \citep{Sanchez:2009aa} and more recently analytical techniques for cultural heritage materials \citep{Csefalvayova:2010aa}, liquid epoxy moulding process \citep{Chan:2011aa}, B-spline surface reconstruction \citep{Galvez:2012aa}, product design \citep{Chan:2012aa} or forecasting cyanotoxins presence in water reservoirs \citep{Garcia-Nieto:2013aa}. These examples not only illustrate the wide spectrum of applications but, additionally, in each one uses, at some point, a \ac{GA}.

\acp{GA} where, arguably, one of the hottest topics of research in the recent decades and with good reason since they outline an optimisation scheme easy to conceptualise and with very broad application. If a nonlinear (or otherwise) model requires parameterization \acp{GA} provide a simple and often effective approach to search for locally optimal parameters. Research related to genetic algorithms abound and spans from the 1950s seminal work of Nils Aall Barricelli \citep{Barricelli:1962aa} in the Institute for Advanced Study of Princeton to today's principal area of study for thousands of researchers, covered in hundreds of conferences, workshops and other meetings. Perhaps the key impulse to \acp{GA} come from John Holland's work and his book ``Adaptation in Natural and Artificial Systems'' \citep{Holland:1975aa}.

One interesting variation of genetic algorithms, named \emph{genetic programming} by John Koza \citep{Koza:1992aa}, proposed the use of \acp{GA} to search the syntactic structure of complex functions. This syntactic structure search is keen to the central ideas of deep learning \citep{Bengio:2009aa,Bengio:2013aa}, a subarea of machine learning actually producing quite promising results (\emph{e.g.} in \cite{Tarlow:2013fk}). It is also related to the work presented in this paper in the sense that, unlike linear models that have a simple structure, $y=\sum_i \beta_i x_i$, nonlinear (in particular polynomial) models pose an additional structure search problem.

@@

\note{Redo the purpose of this work.} This paper describes an extension of the general \ac{EPR} method to find a \emph{regularized} polynomial regression of a given dataset. The optimal regression results from a cost function that accounts for both the \ac{rmse} and a regularization factor to avoid over-fitting.

@@

It turns out that, discarding the computational cost of training, the regression method presented here, \acf{EPRR}, prooves to be quite effective. Indeed, according to the results of common  methods on several real-world datasets it is only systematically out-performed by random forests, an \emph{ensemble} method.
